---
title: "EuroCup Soccer"
author: "Shu Liu"
date: '`r Sys.Date()`'
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## EuroCup Soccer 

This is Data in Motion data analysis challenge #2 More details click here [link](https://d-i-motion.com/lessons/challenge-2-eurocup-soccer/)

### Scenario 

You are a sports data analyst and you have been tasked with summarizing data from the matches from a previous EuroCup. Your manager would like the following questions answered.

### Get the data
Download dataset: Link to [dataset](https://raw.githubusercontent.com/guipsamora/pandas_exercises/master/02_Filtering_%26_Sorting/Euro12/Euro_2012_stats_TEAM.csv)

### Challenge Questions
>
  1. How many teams participated in the Euro2012?
  2. What is the number of columns in the dataset?
  3. View only the columns Team, Yellow Cards and Red Cards and assign them to a dataframe called discipline.
  4. Sort the teams by Red Cards, then to Yellow Cards.
  5. Calculate the mean Yellow Cards given per Team.
  6. Filter teams that scored more than 6 goals.
  7. Select the teams that start with the letter G.
  8. Select the first 7 columns.
  9. Select all columns except the last 3.
  10. Present only the Shooting Accuracy from England, Italy and Russia.

## Steps
### Set up environments
Notes: install package "tidyverse" and "dplyr" for sorting and select.
```{r set up environments}
#an argument repos is added to the function that gives it the web address of the repository.
install.packages("tidyverse", repos = "http://cran.us.r-project.org") 
library(tidyverse)
library(dplyr) #for sorting and select
```
### Load data
Save the data set into local file directory. change working directory to where the file is. load the data into data frame eurocup_soccer.
```{r change working directory}
setwd("C:/Users/liuch/OneDrive/Documents/DataAnalytics/Portfolio/case_study_2")
eurocup_soccer <- read_csv("Euro_2012_stats_TEAM.csv")
```
Now the eurocup_soccer has the data. Let's take a glimpse. 
```{r glimpse the data}
glimpse(eurocup_soccer)
#check is there any missing values.I found 1 missing value.
sum(is.na(eurocup_soccer)) 
```
### Analyze data and answer questions
1. How many teams participated in the Euro2012?
from the glimpse() above we can find out there are **16** teams participated in the Euro2012. 

2. What is the number of columns in the dataset?
From the glimpse() above we can find out there are **35** columns are in the dataset

3. View only the columns Team, Yellow Cards and Red Cards and assign them to a dataframe called discipline.
```{r view Team, Yellow Cards and Red Cards}
descipline <- eurocup_soccer %>% select(Team,`Yellow Cards`,`Red Cards`)
```

4. Sort the teams by Red Cards, then to Yellow Cards.
```{r sort}
descipline %>% arrange(`Yellow Cards`, `Red Cards`)
```

5. Calculate the mean Yellow Cards given per Team.
```{r calculate mean yellow cards }
descipline %>% select(Team, `Yellow Cards`, `Red Cards`) %>% summarise(mean_yellor_cards=mean(`Yellow Cards`))
```
The mean Yellow Cards given per Team is **7.44**

6. Filter teams that scored more than 6 goals.
```{r filter by scores more than 6}
eurocup_soccer %>% filter(Goals>6)
```
Only Germany and Spain have more than 6 goals.

7. Select the teams that start with the letter G.
We can use filter and subtr function
```{r select team start with G}
eurocup_soccer %>% filter(substr(Team,1,1)=="G")
```
Only Germany and Greece start with G.

8. Select the first 7 columns.
```{r select the first 7 columns}
eurocup_soccer %>% select(1:7)
```

9. Select all columns except the last 3.
```{r select all columns but last 3}
eurocup_soccer %>% select(1:(ncol(eurocup_soccer) - 3))
```

10. Present only the Shooting Accuracy from England, Italy and Russia.
```{r show Shooting Accuracy from England,Italy and Russia}
teams_interests<-c("England","Italy","Russia") ## create a vector 
## create a subdataset while Team in the vector teams_interests
subdata_set <- eurocup_soccer %>% filter(Team %in% teams_interests) 
##only select Team and Shooting Accuracy.
subdata_set %>% select(Team, `Shooting Accuracy`) 
```

Notes. This dataset is small and clean. So it's easy to analyze. I also notice if there's space in the column name when we need to use it we need to add **backtitks \`** to quota it, such as **\`Yellow Cards\`**. 